### 2020-10-20 submission/manuscript.Rmd
Create Figures 5 and S6

**How does the removing of rare ASVs impact the probability of falsely detecting a difference between treatment groups?** Observing reduced ability to detect differences between communities when rare ASVs were removed from each sample, I next asked whether removing rare ASVs could lead to falsely claiming that a treatment effect had a significant effect on community diversity and structure. First, for each dataset, I sampled sequences from the null distribution and randomly assigned each sample to one of two treatment groups and determined the number of observed ASVs and OTUs and their Shannon diversity index. Testing at an experiment-wise error rate of 0.05, I expected 5% of the iterations for each dataset to yield a significant test result. Indeed, there was no evidence that removing rare ASVs resulted in an inflated experiment-wise error rate. The average fraction of significant tests did not meaningfully vary from 0.05 across the minimum abundance threshold, dataset, metric of describing sample alpha-diversity, or whether the abundance of ASVs or OTUs were used (**Figure XA**). Similarly, the average fraction of significant tests did not meaningfully vary from 0.05 when using analysis of molecular variance to compare communities using Bray-Curtis distances. Second, I again sampled sequences from the null distribution, but assigned samples to one of two treatment groups based on the number of sequences in each sample. The samples with fewer than the median number of sequences for the dataset were assigned to one group and those with more than the median were assigned to the other. This exaggerated bias has been observed in comparisons of the lung and oral microbiota because of the larger nubmer of non-specific amplicons that can be sequenced from lung samples relative to those in the oral cavity [**REF**]. When rare sequences were not removed, the fraction of significant tests did not differ from 5% for comparing the number of observed taxa (using ASVs or OTUs), their shannon diversity, or Bray-Curtis distances. However, when rare taxa of any frequency were removed, the probability of falsely detecing a difference as signifiant increased with the definition of rarity (**Figure XB**). Across the datasets, the average fraction the average fraction of falsely detecting a difference was when only singletons were removed was greater than 90%. If there is any relationship between the number of sequences and the treatment group, the risk of falsely rejecting the null hypothesis is inflated when researchers use the strategy of removing rare sequences. The most conservative approach is to not remove low abundance sequences.

>>>>

**How does the removing of rare ASVs impact the probability of falsely detecting a difference between treatment groups?** Observing reduced ability to detect differences between communities when rare ASVs were removed from each sample, I next asked whether removing rare ASVs could lead to falsely claiming that a treatment effect had a significant effect on community diversity and structure. First, for each dataset, I sampled sequences from the null distribution and randomly assigned each sample to one of two treatment groups and determined the number of observed ASVs and OTUs and their Shannon diversity index. Testing at an experiment-wise error rate of 0.05, I expected 5% of the iterations for each dataset to yield a significant test result. Indeed, there was no evidence that removing rare ASVs resulted in an inflated experiment-wise error rate. The average fraction of significant tests did not meaningfully vary from 0.05 across the minimum abundance threshold, dataset, metric of describing sample alpha-diversity, or whether the abundance of ASVs or OTUs were used (Figure 5A and S6A). Similarly, the average fraction of significant tests did not meaningfully vary from 0.05 when using analysis of molecular variance to compare communities using Bray-Curtis distances (Figure 5A and S6A). Second, I again sampled sequences from the null distribution, but assigned samples to one of two treatment groups based on the number of sequences in each sample. The samples with fewer than the median number of sequences for the dataset were assigned to one group and those with more than the median were assigned to the other. This exaggerated bias has been observed in comparisons of the lung and oral microbiota because of the larger nubmer of non-specific amplicons that can be sequenced from lung samples relative to those in the oral cavity [**REF**]. When rare sequences were not removed, the fraction of significant tests did not differ from 5% for comparing the number of observed taxa (using ASVs or OTUs), their shannon diversity, or Bray-Curtis distances (Figure 5B and S6B). However, when rare taxa of any frequency were removed, the probability of falsely detecing a difference as signifiant increased with the definition of rarity (Figure 5B and S6B).

Across the datasets, the average fraction the average fraction of falsely detecting a difference was when only singletons were removed was greater than 90%.

If there is any relationship between the number of sequences and the treatment group, the risk of falsely rejecting the null hypothesis is inflated when researchers use the strategy of removing rare sequences. The most conservative approach is to not remove low abundance sequences.
